Modeling memory and taxis to analyze migrations

### 2021-01-06 code/StrategicVsTactical_PartI_v2.Rmd
Minor edits

---
title: "Modeling memory and taxis to analyze migrations"
author: "E. Gurarie, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
theme: cerulean
editor_options:
  chunk_output_type: inline
---

@@

$${\frac{\partial u(x,t)}{\partial t}} = -\varepsilon {\frac{\partial^2 u(x,t)}{\partial x^2}} + \alpha \frac{\partial h(x,t)}{\partial x} + \sum_{i = 1}^n \beta_i \frac{\partial u(x, t - i\tau)}{\partial x}$$

where $u$ represents a population distributed in time and space, $\varepsilon$ is a rate of randomness (diffusion), $\alpha$ is the strength of advection in the direction of the resource $h(x,t)$ gradient), and $\beta_i$ is the strength of advection up the gradient of the *population* distribution in an $i$'th earlier seasonal cycle (year), up to $n \in N$.  The form of $\beta_i$ should be simple, e.g., if the collective memory reaches only one year back then $n = 1$ and $\beta$ is a single constant. Alternatively, a form of memory might discounts older "traditions" via $\beta_i = \beta \exp(-i/\gamma)$, where $\gamma$ represents a generational scale of memory.

@@

```{r}
ForagingMemoryModel <- function(t, y, parms, resource, pop_lag){
  tran.1D(C = y, D = parms["epsilon"], 
          flux.up = 0, flux.down = 0, 
          v = parms["alpha"] * diff(resource)/parms["dx"] +  
            parms["gamma"] * diff(y)/parms["dx"] +
            parms["beta"] * diff(pop_lag)/parms["dx"], 
          dx = parms["dx"])
}

ForagingMemoryModel <- function(t, y, parms, resource, pop_lag){
  tran.1D(C = y, D = parms["epsilon"], 
          flux.up = 0, flux.down = 0, 
          v = parms["alpha"] * diff(resource)/parms["dx"] +  
            parms["gamma"] * diff(y)/parms["dx"] +
            parms["beta"] * diff(pop_lag)/parms["dx"], 
          dx = parms["dx"])
}

```{r memoryOnly, fig.height = 5, fig.width = 10}
dx <- .5
parameters <- c(epsilon = 5, alpha = 0, beta = 5e2, gamma = 20, lambda = 0, dx = dx)

>>>>

---
title: "Modeling memory and taxis to analyze migrations"
author: "E. Gurarie, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
theme: cerulean
editor_options:
  chunk_output_type: console
---

@@

$${\frac{\partial u(x,t)}{\partial t}} = -\varepsilon {\frac{\partial^2 u(x,t)}{\partial x^2}} + \alpha \frac{\partial h(x,t)}{\partial x} + \sum_{i = 0}^n \beta_i \frac{\partial u(x, t - i\tau)}{\partial x}$$

where $u$ represents a population distributed in time and space, $\varepsilon$ is a rate of randomness (diffusion), $\alpha$ is the strength of advection in the direction of the resource $h(x,t)$ gradient), and $\beta_i$ is the strength of advection up the gradient of the *population* distribution in an $i$'th earlier seasonal cycle (year), up to $n \in N$.

Note that $\beta_0$ is similar to an index of "cohesion", essentially encouraging the population to stay close together.  Higher terms of $\beta_i$ should be simple, e.g., if the collective memory reaches only one year back then $n = 1$ and $\beta_i$ is a single constant. Alternatively, a form of memory might discounts older "traditions" via $\beta_i = \beta \exp(-i/\gamma)$, where $\gamma$ represents a generational scale of memory.

@@

```{r}
ForagingMemoryModel <- function(t, y, parms, resource, pop_lag){
  tran.1D(C = y, D = parms["epsilon"], 
          flux.up = 0, flux.down = 0, 
          v = parms["alpha"] * diff(resource)/parms["dx"] +  
            parms["beta0"] * diff(y)/parms["dx"] +
            parms["beta1"] * diff(pop_lag)/parms["dx"], 
          dx = parms["dx"])
}
```

```{r memoryOnly, fig.height = 5, fig.width = 10}
dx <- .5
parameters <- c(epsilon = 5, alpha = 0, 
                beta0 = 20, 
                beta1 = 5e2, 
                lambda = 0, dx = dx)
